### 2017-05-02 handout.tex
handout for human subjects

\documentclass{article}
\usepackage{tikz}
\usepackage[margin=0.1in]{geometry}
\begin{document}
\centering\textbf{CoCoSci NIPS Data Blitz!}

Draw diagrams on a grid. You can draw rectangles, lines, arrows,
dashed lines, and circles of radius 1.  Your drawings don't have to be
perfect, but they should be good enough for a human to understand. On
the last grid you can draw whatever diagram you want.

\centering
\begin{tikzpicture}[scale = 0.5]
  \node[align = right] at (5,14) {Here is an example diagram:};
  \draw[step=1cm,gray,ultra thin] (0,0) grid (13,13);
  \draw[ultra thick] (3,3) circle (1cm);
  \draw[ultra thick] (3,8) circle (1cm);
  \draw[line width = 0.1cm,dashed] (3,4) -- (3,7);
  \draw[ultra thick] (8,3) rectangle (13,10);
  \draw[ultra thick] (9,4) rectangle (12,9);
  \draw[ultra thick,-{>[scale = 1.5]}]  (4,3) -- (8,4);
  \draw[ultra thick,-{>[scale = 1.5]}]  (4,3+5) -- (8,4+5);

    \node[align = right] at (5 + 19,14) {Can you copy the diagram on to this grid?};
    \draw[step=1cm,gray,ultra thin] (17,0) grid (13+17,13);

  \node[align = right] at (5,14 - 16) {Come up with a new diagram:};
  \draw[step=1cm,gray,ultra thin] (0,0-16) grid (13,13 - 16);

  \node[align = right] at (5 + 17,14 - 16) {Come up with a new diagram:};
  \draw[step=1cm,gray,ultra thin] (17,0-16) grid (13 + 17,13 - 16);

  \node[align = right] at (5,14 - 16 - 16) {Come up with a new diagram:};
  \draw[step=1cm,gray,ultra thin] (0,0-16 - 16) grid (13,13 - 16 - 16);

  \node[align = center] at (8 + 17,14 - 16 - 16) {Diagram anything you want\\(stars, text, ellipses, spirals, thought bubbles, fractals, flowcharts, ...)};
  \draw[step=1cm,gray,ultra thin] (17,0-16 - 16) grid (13 + 17,13 - 16 - 16);
\end{tikzpicture}

\end{document}