### 2017-05-13 extrapolations.tex
preliminary extrapolation results

            \begin{tabular}{ll}
    \includegraphics[width = 5cm]{../TikZ/drawings/expert-17.png}&
    \includegraphics[width = 5cm]{../TikZ/extrapolations/expert-17-extrapolation.png}
    \end{tabular}        
            \\

            \begin{tabular}{ll}
    \includegraphics[width = 5cm]{../TikZ/drawings/expert-21.png}&
    \includegraphics[width = 5cm]{../TikZ/extrapolations/expert-21-extrapolation.png}
    \end{tabular}        
            \\

            \begin{tabular}{ll}
    \includegraphics[width = 5cm]{../TikZ/drawings/expert-7.png}&
    \includegraphics[width = 5cm]{../TikZ/extrapolations/expert-7-extrapolation.png}
    \end{tabular}        
            \\

            \begin{tabular}{ll}
    \includegraphics[width = 5cm]{../TikZ/drawings/expert-36.png}&
    \includegraphics[width = 5cm]{../TikZ/extrapolations/expert-36-extrapolation.png}
    \end{tabular}        
            \\

            \begin{tabular}{ll}
    \includegraphics[width = 5cm]{../TikZ/drawings/expert-58.png}&
    \includegraphics[width = 5cm]{../TikZ/extrapolations/expert-58-extrapolation.png}
    \end{tabular}        
            \\

            \begin{tabular}{ll}
    \includegraphics[width = 5cm]{../TikZ/drawings/expert-22.png}&
    \includegraphics[width = 5cm]{../TikZ/extrapolations/expert-22-extrapolation.png}
    \end{tabular}        
            \\

            \begin{tabular}{ll}
    \includegraphics[width = 5cm]{../TikZ/drawings/expert-14.png}&
    \includegraphics[width = 5cm]{../TikZ/extrapolations/expert-14-extrapolation.png}
    \end{tabular}        
            \\

            \begin{tabular}{ll}
    \includegraphics[width = 5cm]{../TikZ/drawings/expert-68.png}&
    \includegraphics[width = 5cm]{../TikZ/extrapolations/expert-68-extrapolation.png}
    \end{tabular}        
            \\

            \begin{tabular}{ll}
    \includegraphics[width = 5cm]{../TikZ/drawings/expert-12.png}&
    \includegraphics[width = 5cm]{../TikZ/extrapolations/expert-12-extrapolation.png}
    \end{tabular}        
            \\

            \begin{tabular}{ll}
    \includegraphics[width = 5cm]{../TikZ/drawings/expert-35.png}&
    \includegraphics[width = 5cm]{../TikZ/extrapolations/expert-35-extrapolation.png}
    \end{tabular}        
            \\

            \begin{tabular}{ll}
    \includegraphics[width = 5cm]{../TikZ/drawings/expert-31.png}&
    \includegraphics[width = 5cm]{../TikZ/extrapolations/expert-31-extrapolation.png}
    \end{tabular}        
            \\

            \begin{tabular}{ll}
    \includegraphics[width = 5cm]{../TikZ/drawings/expert-37.png}&
    \includegraphics[width = 5cm]{../TikZ/extrapolations/expert-37-extrapolation.png}
    \end{tabular}        
            \\

            \begin{tabular}{ll}
    \includegraphics[width = 5cm]{../TikZ/drawings/expert-29.png}&
    \includegraphics[width = 5cm]{../TikZ/extrapolations/expert-29-extrapolation.png}
    \end{tabular}        
            \\

            \begin{tabular}{ll}
    \includegraphics[width = 5cm]{../TikZ/drawings/expert-57.png}&
    \includegraphics[width = 5cm]{../TikZ/extrapolations/expert-57-extrapolation.png}
    \end{tabular}        
            \\

            \begin{tabular}{ll}
    \includegraphics[width = 5cm]{../TikZ/drawings/expert-23.png}&
    \includegraphics[width = 5cm]{../TikZ/extrapolations/expert-23-extrapolation.png}
    \end{tabular}        
            \\

            \begin{tabular}{ll}
    \includegraphics[width = 5cm]{../TikZ/drawings/expert-60.png}&
    \includegraphics[width = 5cm]{../TikZ/extrapolations/expert-60-extrapolation.png}
    \end{tabular}        
            \\

            \begin{tabular}{ll}
    \includegraphics[width = 5cm]{../TikZ/drawings/expert-16.png}&
    \includegraphics[width = 5cm]{../TikZ/extrapolations/expert-16-extrapolation.png}
    \end{tabular}        
            \\

            \begin{tabular}{ll}
    \includegraphics[width = 5cm]{../TikZ/drawings/expert-5.png}&
    \includegraphics[width = 5cm]{../TikZ/extrapolations/expert-5-extrapolation.png}
    \end{tabular}        
            \\

            \begin{tabular}{ll}
    \includegraphics[width = 5cm]{../TikZ/drawings/expert-50.png}&
    \includegraphics[width = 5cm]{../TikZ/extrapolations/expert-50-extrapolation.png}
    \end{tabular}        
            \\

            \begin{tabular}{ll}
    \includegraphics[width = 5cm]{../TikZ/drawings/expert-6.png}&
    \includegraphics[width = 5cm]{../TikZ/extrapolations/expert-6-extrapolation.png}
    \end{tabular}