### 2017-05-19 supplement.tex
attention citation; improvements to supplement

\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
% \PassOptionsToPackage{numbers, compress}{natbib}
% before loading nips_2016
%
% to avoid loading the natbib package, add option nonatbib:
% \usepackage[nonatbib]{nips_2016}

\usepackage{dsfont}

% to compile a camera-ready version, add the [final] option, e.g.:
\usepackage{nips_2017}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

\usepackage{listings}
\usepackage{amsthm}
% use Times
\usepackage{times}
% For figures
\usepackage{graphicx} % more modern
%\usepackage{epsfig} % less modern
\usepackage{subfig} 
\usepackage{fancyvrb}

\usepackage{caption}
\usepackage{subcaption}

\fvset{fontsize=\footnotesize}

\usepackage{amssymb}
\usepackage{listings}
\usepackage{wrapfig}
\usepackage{tabularx}

\usepackage{verbatim}
 \usepackage{booktabs}
 % For algorithms
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{tikz}
% As of 2011, we use the hyperref package to produce hyperlinks in the
% resulting PDF.  If this breaks your system, please commend out the
% following usepackage line and replace \usepackage{icml2016} with
% \usepackage[nohyperref]{icml2016} above.
\usepackage{amsmath}
\usepackage{hyperref}
\DeclareMathOperator*{\argmin}{arg\,min} % thin space, limits underneath in displays
\DeclareMathOperator{\argmin}{argmin} % no space, limits underneath in displays

% Packages hyperref and algorithmic misbehave sometimes.  We can fix
% this with the following command.

\newcommand{\Expect}{\mathds{E}} %{{\rm I\kern-.3em E}}
\newcommand{\probability}{\mathds{P}} %{{\rm I\kern-.3em P}}

\newtheorem{proposition}{Proposition}

\newcommand\modt{\stackrel{\mathclap{\normalfont 2}}{\equiv}}

\title{Supplement to: Inferring Graphics Programs from Images}

% The \author macro works with any number of authors. There are two
% commands used to separate the names and addresses of multiple
% authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to
% break the lines. Using \AND forces a line break at that point. So,
% if LaTeX puts 3 of 4 authors names on the first line, and the last
% on the second line, try using \AND instead of \And before the third
% author name.

\author{
Kevin Ellis \\
Brain and Cognitive Sciences\\
MIT\\
%Pittsburgh, PA 15213 \\
\texttt{ellisk@mit.edu} \\
Daniel Ritchie\\
Department of Computer Science\\
Brown
\And
Armando Solar-Lezama \\
  CSAIL\\
MIT \\
\texttt{asolar@csail.mit.edu} \\
\And
Joshua B. Tenenbaum \\
Brain and Cognitive Sciences\\
MIT\\
\texttt{jbt@mit.edu} \\
}

\begin{document}
% \nipsfinalcopy is no longer used

\maketitle

\section{Neural network architecture}

\subsection{Convolutional network}
The convolutional network takes as input 2 $256\times 256$ images
represented as a $2\times 256\times 256\times$ volume. These are
passed through two layers of convolutions separated by ReLU
nonlinearities and max pooling:
\begin{itemize}
\item Layer 1: 20 $8\times 8$ convolutions, 2 $16\times 4$ convolutions, 2 $4\times 16$ convolutions. Followed by $8\times 8$ pooling with a stride size of 4.
\item Layer 2: 10 $8\times 8$ convolutions. Followed by $4\times 4$ pooling with a stride size of 4.
\end{itemize}
Training takes a little bit less than a day on a Nvidia TitanX GPU.
The network was trained on $10^5$ synthetic examples.

\subsection{Autoregressive decoding of drawing commands}

Given the image features $f$, we predict the first token using logistic regression:
\begin{equation}
  \probability [T_1]\propto W_{T_1}f
\end{equation}
where $W_{T_1}$ is a learned weight matrix.

Subsequent tokens are predicted as:
\begin{equation}
  \probability [T_n|T_{1:(n - 1)}]\propto \text{MLP}_{T_1,n}(I \otimes \bigotimes_{j < n} \text{oneHot}(T_j))
\end{equation}
Thus each token of each drawing primitive has its own learned MLP.
For predicting the coordinates of lines we found that using 32 hidden nodes with sigmoid activations worked well;
for other tokens the MLP's are just logistic regression (no hidden nodes).

\subsection{A learned likelihood surrogate}

Our architecture for
$L_{\text{learned}}(\text{render}(T_1)|\text{render}(T_2))$ has the
same series of convolutions as the network that predicts the next
drawing command. We train it to predict two scalars: $|T_1 - T_2|$ and
$|T_2 - T_1|$.  These predictions are made using linear regression
from the image features followed by a ReLU nonlinearity; this
nonlinearity makes sense because the predictions can never be negative
but could be arbitrarily large positive numbers.

We train this network by sampling random synthetic scenes for $T_1$,
and then perturbing them in small ways to produce $T_2$.
We minimize the squared loss between the network's prediction and the ground truth symmetric differences.
$T_1$ is rendered in a ``simulated hand drawing'' style which we describe next.

\section{Simulating hand drawings}

We introduce noise into the \LaTeX~rendering process by:

\begin{itemize}
\item Rescaling the image intensity by a factor chosen uniformly at random from $[0.5,1.5]$
\item Translating the image by $\pm 3$ pixels chosen uniformly random
\item Rendering the \LaTeX~using the \verb|pencildraw| style,
  which adds random perturbations to the paths drawn by \LaTeX in a way designed to resemble a pencil.
\item Randomly perturbing the positions and sizes of primitive  \LaTeX drawing commands
\end{itemize}

\section{Likelihood surrogate for synthetic data}
For synthetic data (e.g., \LaTeX~output)
it is relatively straightforward to engineer an adequate distance measure between images,
because it is possible for the system to discover
drawing commands that exactly match the pixels in the target image.
We use:
\begin{equation}
  -\log L(I_1|I_2) = \sum_{1\leq x\leq 256} \sum_{1\leq y\leq 256} |I_1[x,y] - I_2[x,y]|\begin{cases}
    \alpha\text{, if }$I_1[x,y] > I_2[x,y]$\\
    \beta\text{, if }$I_1[x,y] <  I_2[x,y]$\\
        0\text{, if }$I_1[x,y] = I_2[x,y]$\\
    \end{cases}
  \end{equation}
where $\alpha$, $\beta$ are constants that control the trade-off
between preferring to explain the pixels in the image (at the expense
of having extraneous pixels) and not predicting pixels where they
don't exist (at the expense of leaving some pixels unexplained). Because our sampling procedure incrementally constructs the scene part-by-part,
we want $\alpha > \beta$.
That is, it is preferable to leave some pixels unexplained;
for once a particle in SMC adds a drawing primitive to its trace that is not actually in the latent scene,
it can never recover from this error.
In our experiments on synthetic data we used $\alpha = 0.8$ and $\beta = 0.04$. 
\section{Generating synthetic training data}

We generated synthetic training data for the neural network by
sampling \LaTeX~code according to the following generative process:
First,
the number of objects in the scene are sampled uniformly from 1 to 8.
For each object we uniformly sample its identity (circle, rectangle, or line).
Then we sample the parameters of the circles,
than the parameters of the rectangles,
and finally the parameters of the lines;
this has the effect of teaching the network to first draw the circles in the scene,
then the rectangles,
and finally the lines.
We furthermore put the circle (respectively, rectangle and line) drawing commands in order by
left-to-right, bottom-to-top;
thus the training data enforces a canonical order in which to draw any scene.

To make the training data look more like naturally occurring figures,
we put a Chinese restaurant process prior~\cite{gershman2012tutorial} over
the values of the X and Y coordinates that occur in the execution trace.
This encourages reuse of coordinate values,
and so produces training data that tends to have parts that are nicely aligned.

In the synthetic training data we excluded any sampled scenes that had
overlapping drawing commands.  As shown in the main paper, the network
is then able to generalize to scenes with, for example, intersecting
lines or lines that penetrate a rectangle.

When sampling the endpoints of a line,
we biased the sampling process so that it would
be more likely to start an endpoint
along one of the sides of a rectangle or at the boundary of a circle.
If $n$ is the number of points either along the side of a rectangle or at the boundary of a circle,
we would sample an arbitrary endpoint with probability $\frac{2}{2 + n}$
and sample one of the ``attaching'' endpoints with probability $\frac{1}{2 + n}$.

See figure~\ref{exampleTrainingData} for examples of the kinds of scenes that the network is trained on.
\begin{figure}
  \begin{minipage}[t]{2.2cm}\fbox{\includegraphics[width=2cm]{figures/randomScene-58-7.png}}\end{minipage}
  \begin{minipage}[t]{2.2cm}\fbox{\includegraphics[width=2cm]{figures/randomScene-33-5.png}}\end{minipage}
  \begin{minipage}[t]{2.2cm}\fbox{\includegraphics[width=2cm]{figures/randomScene-4-4.png}}\end{minipage}
  \begin{minipage}[t]{2.2cm}\fbox{\includegraphics[width=2cm]{figures/randomScene-26-7.png}}\end{minipage}
      \begin{minipage}[t]{2.2cm}\fbox{\includegraphics[width=2cm]{figures/randomScene-25-4.png}}\end{minipage}\\
      \begin{minipage}[t]{2.2cm}\fbox{\includegraphics[width=2cm]{figures/randomScene-22-3.png}}\end{minipage}
      \begin{minipage}[t]{2.2cm}\fbox{\includegraphics[width=2cm]{figures/randomScene-20-7.png}}\end{minipage}
      \begin{minipage}[t]{2.2cm}\fbox{\includegraphics[width=2cm]{figures/randomScene-19-3.png}}\end{minipage}
      \begin{minipage}[t]{2.2cm}\fbox{\includegraphics[width=2cm]{figures/randomScene-17-7.png}}\end{minipage}
      \begin{minipage}[t]{2.2cm}\fbox{\includegraphics[width=2cm]{figures/randomScene-1-4.png}}\end{minipage}
      \caption{Example synthetic training data}\label{exampleTrainingData}
\end{figure}

For readers wishing to generate their own synthetic training sets,
we refer them to our source code at:~\url{http://www.redactedForAnonymousReview.com}.

\section{The cost function for programs}

We seek the minimum cost program which evaluates to (produces the drawing primitives in) an execution trace $T$:
\begin{equation}
  \text{program}(T) = \argmin_{\substack{p\in \text{DSL}\\p \text{ evaluates to } T}} \text{cost}(p)\label{programObjective}
\end{equation}

Programs incur a cost of 1 for each command (primitive drawing action,
loop, or reflection).  They incur a cost of $\frac{1}{3}$ for each
unique coefficient they use in a linear transformation beyond the
first coefficient. This encourages reuse of coefficients, which leads
to code that has translational symmetry; rather than provide a
translational symmetry operator as we did with reflection, we modify
what is effectively a prior over the space of program so that it tends
to produce programs that have this symmetry.

Programs also incur a cost of 1 for having loops of constant length 2;
otherwise there is often no pressure from the cost function to explain
a repetition of length 2 as being a reflection rather a loop.

%% \section{Neural networks for guiding SMC}

%% Let $L(\cdot | \cdot):\text{image}^2\to \mathcal{R}$ be our likelihood
%% function: it takes two images, an observed target image and a
%% hypothesized program output, and gives the likelihood of the observed
%% image conditioned on the program output. We want to sample from:
%% \begin{equation}
%% \probability [p|x]  \propto L(x | \text{render}(p)) \probability [p]
%% \end{equation}
%% where $\probability [p]$ is the prior probability of program $p$, and $x$ is the observed image.

%% Let $p$ be a program with $L$ lines, which we will write as $p = (p_1,p_2,\cdots,p_L)$. Assume the prior factors into:
%% \begin{equation}
%%   \probability [p]\propto \prod_{l\leq L}\probability [p_l]
%% \end{equation}
%% Define the distribution $q_L(\cdot)$, which happens to be proportional to the above posterior:
%% \begin{equation}
%%   q_L(p_1,p_2,\cdots,p_{L - 1},p_L)\propto q_{L - 1}(p_1,p_2,\cdots,p_{L - 1})\times \frac{L(x | \text{render}(p_1,p_2,\cdots,p_{L - 1},p_L))}{L(x | \text{render}(p_1,p_2,\cdots,p_{L - 1}))}\times\probability [p_L]
%% \end{equation}
%% Now suppose we have some samples from $q_{L - 1}(\cdot)$, and that we
%% then sample a $p_L$ from a distribution proportional to $\frac{L(x |
%%   \text{render}(p_1,p_2,\cdots,p_{L - 1},p_L))}{L(x |
%%   \text{render}(p_1,p_2,\cdots,p_{L - 1}))}\times\probability [p_L]$.
%% The resulting programs $p$ are distributed according to $q_L$, and so
%% are also distributed according to $\probability [p|x]$.

%% How do we sample $p_L$ from a distribution proportional to $\frac{L(x
%%   | \text{render}(p_1,p_2,\cdots,p_{L - 1},p_L))}{L(x |
%%   \text{render}(p_1,p_2,\cdots,p_{L - 1}))}\times\probability [p_L]$?
%% We have a neural network that takes as input the target image $x$ and
%% the program so far, and produces a distribution over next lines of
%% code ($p_L$).  We write $\text{NN}(p_L | p_1,\cdots,p_{L - 1};x)$ for
%% the distribution output by the neural network. So we can sample from NN and then weight the samples by:
%% \begin{equation}
%%   w(p_L) = \frac{\probability [p_L]}{\text{NN}(p_L | p_1,\cdots,p_{L - 1};x)}\times \frac{L(x | \text{render}(p_1,p_2,\cdots,p_{L - 1},p_L))}{L(x | \text{render}(p_1,p_2,\cdots,p_{L - 1}))}
%% \end{equation}
%% Then we can resample from these now weighted samples to get a new
%% population of particles (here programs are particles), where each
%% program now has $L$ lines instead of $L - 1$.

%% This procedure can be seen as a particle filter, where each successive
%% latent variable is another line of code, and the emission
%% probabilities are successive ratios of likelihoods under $L(\cdot |
%% \cdot)$.

%%   \begin{algorithm}[tb]
%%    \caption{Neurally guided SMC}
%%    \label{guideAlgorithm}
%% \begin{algorithmic}
%%   \STATE {\bfseries Input:} Neural network NN, beam size $N$, maximum length $L$, target image $x$
%%   \STATE {\bfseries Output:} Samples of the program trace
%%   \STATE Set $B_0 = \{\text{empty program}\}$
%%   \FOR{$1\leq l\leq L$}
%%   \FOR{$1\leq n\leq N$}
%%   \STATE{ $p_n\sim \text{Uniform}(B_{l - 1})$}
%%   \STATE{ $p'_{n}\sim \text{NN}(\text{render}(p),x)$}
%%   \STATE{ Define $r_n = p'_n\cdot p_n$}
%%   \STATE{ Set $\tilde{w}(r_n) = \frac{L(x|r_n)}{L(x|p_n)}\times\frac{\probability [p'_n]}{\probability [p'_n = \text{NN}(\text{render}(p),x)]}$}
%%   \ENDFOR
%%   \STATE{ Define $w(p) = \frac{\tilde{w}(p)}{\sum_{p'}\tilde{w}(p')}$}
%%   \STATE{ Set $B_l$ to be $N$ samples from $r_n$ distributed according to $w(\cdot)$}
%%   \ENDFOR
%%   \STATE {\bfseries return} $\{p : p\in B_{l\leq L}, p \text{ is finished}\}$
%% \end{algorithmic}
%%   \end{algorithm}

\section{Full results on drawings data set}

Below we show our full data set of drawings. The leftmost column is a hand drawing. The middle column is a rendering of the most likely trace discovered by the neurally guided SMC sampling scheme. The rightmost column is the program we synthesized from a ground truth execution trace of the drawing.
\input{synthesizerOutputs.tex}

\end{document}